\Appendix{In context stimuli}\label{Appendix::InContext}

\section{English --- Syntax Nonspecific --- Cognates}
The lion that sitting on the field saw the antelope while it was eating grass by the pond.	\\	\\
The farmer who labored intensively in the field hated the artist for the leisurely life he led.	\\	\\
The teachers who went to the event cheered on the athlete during his final race.	\\	\\
The woman who had made the call to the police spoke to the authorities about the burglary when they arrived.	\\	\\
The baker who works near the church has a baby who is four years old who can already play the piano.	\\	\\
The girl who I met at the store is a biologist who makes pharmaceuticals.	\\	\\
The girl who always forgets where she leaves her keys went to the store to buy some broccoli for the dish that she was preparing for dinner.	\\	\\
The man who was returning from the dinner drove the bus with the teachers from his school.	\\	\\
The postman who works for the city cut the cable that hung from the window because it caused a safety hazard.	\\	\\
The lawyer who studied at Harvard ordered the camera on the Internet because it was less expensive.	\\	\\
The historian who was sitting in the park knew all the capitals of the African countries.	\\	\\
The psychologist who studies memory loves it when her participants bring her caramels as a present.	\\	\\
The butler who works in the mansion called the carpenter to fix the main door of the guest house.	\\	\\
The mason who was cousin to the king built a cathedral with recycled materials.	\\	\\
The dalmatian that had very floppy ears ate the cereal the child spilled onto the floor.	\\	\\
The boy who is in third grade drew a circle with his blue pencil and then he cut it and glued it on his book.	\\	\\
The woman who had worked at a company that manufactures paper saw her client in the mall.	\\	\\
The actress who was on the balcony killed a cobra with a shovel that she grabbed from the shed.	\\	\\
The thief who robbed the bank drank some coffee before the robbery and then told his friends what he did.	\\	\\
The bookseller who opened the new store on the corner invited his colleagues to the inauguration.	\\	\\
The telegrapher who worked in the air base warned the colonel of the attack that the enemies were planning.	\\	\\
The actress who studied in Madrid had dinner with the composer from San Francisco.	\\	\\
The teller who works at the bank burned the computer on her desk with her cigarette.	\\	\\
The assistant who teaches at the school designed the cube for the Physics class.	\\	\\
The congressman who lives in Minnesota argued that his decision was the right one given the circumstances.	\\	\\
The writer who was trying to get ideas for his novel listened to the detective during the investigation.	\\	\\
The businessman who had invested a lot of money fired the director of the company  last month.	\\	\\
The soldier who was anxious to prove himself set off the dynamite too early and almost jeopardized the mission.	\\	\\
The accountant who has three children called the engineer on his wedding day to congratulate him.	\\	\\
The man who loves animals brought home a kitten for his family without first telling his wife.	\\	\\
The hairdresser who bought a car opened the garage of her house with her remote control.	\\	\\
The king who governed the country invited the general to eat shrimp and drink a beer.	\\	\\
The firemen who arrived at the scene of the accident helped the giraffe in danger at the San Diego Zoo.	\\	\\
The gardener who was hired by the school prepared a hamburger to eat with his friends.	\\	\\
The archaeologist who has dedicated his life to science inaugurated the institute of Anatomic Research.	\\	\\
My father who has traveled the world had never seen a kangaroo until his trip to Australia.	\\	\\
The people who founded the United Nations were members of a hopeful generation of individuals.	\\	\\
The hostage who was trapped in the bank called the official while the thief was distracted.	\\	\\
The nurse who will be working with the cardiologist pushed the organizer out the door.	\\	\\
The specialist who is very famous spoke to the patient about various treatments to cure his sickness.	\\	\\
The dentist who died last week had bought a pipe in Turkey for his collection.	\\	\\
The man who sailed across the ocean in his ship killed the pirate with a knife.	\\	\\
The waitress who smokes two packs of cigarettes a day carries a pistol in her handbag.	\\	\\
The politician who lives in Washington bought a plant for his wife because she likes them very much.	\\	\\
The woman who opened the cabinet threw a plate at her husband because she was angry at him.	\\	\\
The news anchor who had been promoted recently encouraged the presenter to act naturally on the air.	\\	\\
The diplomat who travels to many countries asked the president whether she could become a spy.	\\	\\
The decorator who writes for the magazine predicted that the problems with the paint would be hard to solve.	\\	\\
The taxi driver who was parked at the corner of the bakery took the professor to her house.	\\	\\
The explorer who was interviewed yesterday said that the project required more funds for its completion.	\\	\\
The man who organized the event found a rat in the pantry of his kitchen last week.	\\	\\
The governor's sister who helped him during his campaign picked up the letters from the receptionist before leaving.	\\	\\
The skier who won the medal in the Olympics saw the reporter during his jump.	\\	\\
The baseball that was signed by a famous player landed on the sofa after it fell off the cabinet.	\\	\\
The lady who lives in our building left the soup on the window so that it would cool off.	\\	\\
The dancer who was at the exhibit will buy a statue in Berlin during her next visit.	\\	\\
The trainer who works for the school believes that his strategy will give good results.	\\	\\
The woman who attended summer camp as a child became the best student in her class.	\\	\\
The soldier who had fought in the battle sewed the sweater in his camping tent.	\\	\\
The pacifist who had organized the demonstration painted the tank with peace signs.	\\	\\
The caretaker who began to work yesterday cleaned the telephone with water and soap to disinfect it.	\\	\\
The spy who turned over the data interrogated the tourists in Cancun to obtain more information about the case.	\\	\\
The kids who were at the supermarket called the vendor through the window.	\\	\\
The actor who lived in Chile bought a violin for her daughter's birthday from a famous store.	\\	\\


\section{English --- Syntax Nonspecific --- Noncognates}
The audience who attended the fashion show recognized the beauty of the model from Romania.	\\	\\
The nun who had lost her mind attacked the beggar in the kitchen on Thanksgiving day.	\\	\\
The dog that was very sweet watched the blind to keep him out of harm's way.	\\	\\
The man who lives down the street has three boys who follow him everywhere he goes.	\\	\\
The businessman who worked obsessively around the clock watched the building while it was being demolished.	\\	\\
The historian who was recording a TV show investigated the cave before starting the recording.	\\	\\
The firefighter who was in danger inside the building remembered his childhood as he saved the newborn child.	\\	\\
The lawyer who dealt with the purchase of the building sent the deed to the city hall from the Post Office.	\\	\\
The dietitian who celebrated his birthday with his friends prepared a dessert for the dinner.	\\	\\
The tenant who loves animals always combed his dog with a natural bristle brush.	\\	\\
The agent who hated physical activity called the elevator to go the first floor.	\\	\\
The story that the boy was reading involved a group of elves who where trying to save the world from evil trolls.	\\	\\
The driver who had the day off for the break called the fisherwoman to go out.	\\	\\
The pie maker who owned a pastry shop in the city threw the flour on the main street to protest.	\\	\\
The lady who was doing the dishes removed the foam from her hands before answering the door.	\\	\\
The country that was formed ten years ago prohibited foreigners from becoming residents.	\\	\\
The grandmother who had won the lottery bought a freezer for her new house.	\\	\\
The woman who was hired by the company used to be a friend of the CEO who lived in Panama.	\\	\\
The zoologist who conducted experiments with monkeys stopped their growth with the compound that he had prepared.	\\	\\
The woman who always keeps a pristine house persuaded her guests to take off their shoes when they entered.	\\	\\
The girl who is an Olympic gymnast does not want to cut her hair despite the fact that it is too long.	\\	\\
The stylist who owns a store downtown sells hairspray and other products for hair.	\\	\\
The assistant who is in charge of preparing pizzas learned the proper handling of the machine to make dough.	\\	\\
The astronomer who worked at the observatory in Houston had a hole in his jacket.	\\	\\
The boy who wants to be a vet wanted to buy a horse with the money from his savings account.	\\	\\
The jockey who won the last race of the Kentucky Derby bought iron for his stable.	\\	\\
The painter who worked for my father in the mornings took the ladder out from the van.	\\	\\
My son who eats meat said that his friend bought lamb for the picnic.	\\	\\
The woman who needed a book stopped the librarian to ask her a question.	\\	\\
The pilot who threw a celebration at his house bought the best loin and ham for sale at the store.	\\	\\
The minister who came back from his holiday in Italy fired the manager for being incompetent.	\\	\\
The neighbor who lived on the sixth floor of the building decorated the neighborhood for the festivities.	\\	\\
The boy who lives near my house bought a notebook to take notes on a field trip.	\\	\\
The guide who visited  several cities in Australia prepared an ostrich in the kitchen.	\\	\\
The writer who arrived in Madrid yesterday saw the parade from the balcony with his friends.	\\	\\
The writer who was concentrating on his new book didn't hear the party in the park.	\\	\\
The little girl who was wearing her favorite dress ate the peaches in the car on her way back home.	\\	\\
The girl who wants to be an artist received a prize for her performance in the school play.	\\	\\
The man who prepares jams pureed the plums and cooked them on the stove.	\\	\\
The cook who prepares stews bought a potato at the market for dinner.	\\	\\
The agent who rented a car at the airport fixed the printer for the company.	\\	\\
The swimmer who ate cakes and cookies received a punishment from her coach before the competition.	\\	\\
The carrot that the farmer had picked earlier that day hit the puppy causing it to yelp.	\\	\\
The police officer who visited the park arrested the rangers who were suspected of robbing a bank.	\\	\\
The man who likes to exercise decided to become a referee after seeing a soccer match.	\\	\\
The trainer who screamed loudly during the game tore the report with his notes.	\\	\\
The sister who is very meticulous bought a ribbon for the bride's veil.	\\	\\
The nurse who had a free hour after her lunch cleaned the scale in the laboratory.	\\	\\
The teacher who wanted to decorate the bedroom of her new house bought a screen on Friday.	\\	\\
The janitor who needed clothes for her job ordered some socks on the Internet.	\\	\\
The priest who preaches every Sunday saved the soul of every sinner last week.	\\	\\
The man who had taken a course on survival skills created a spark to start the fire.	\\	\\
The teenager who was in love with a friend waited for the arrival of the spring to express his love for her.	\\	\\
The plumber who had three kids at the university prepared a steak with vegetables.	\\	\\
The doctor who drank too many beers at the bar hurt his stepbrother in a fight.	\\	\\
The student who participated in an experiment returned the survey for the study in an envelope.	\\	\\
The business man who dressed very well bought a tie in a store next to his house.	\\	\\
The waiter who ran every morning arrived at the tower in less than an hour.	\\	\\
The actress who loved her job more than anything else invited the traveler for a glass of wine.	\\	\\
The juror who had been listening to the case for three days straight faced the wall when he declared that he needed a rest.	\\	\\
The student who had many chores completed the wash with fabric softener before going back home.	\\	\\
The man who is very intelligent but messy cleaned his workroom before his friend arrived.	\\	\\
The wrestler who defeated the world champion covered his wound with a bandage after the fight.	\\	\\
The goalkeeper who played soccer for the school sewed the zipper of his pants before the game.	\\	\\

\section{English --- Syntax Specific --- Cognates}
The veterinarian injected the antibiotics that she had in her cabinet to the antelope after it began to shake uncontrollably.	\\	\\
The rich businessman gave the photograph that he had taken of his wife to the artist in order to paint her portrait.	\\	\\
The nurse gave the medicine that she dissolved in water to the athlete with a headache.	\\	\\
The suspect gave the knife that he had been hiding to the authorities when he decided to turn himself in.	\\	\\
The uncle gave the crib that he had restored a few months ago to the baby for her birthday.	\\	\\
The police gave the blood sample that they had found at the crime scene to the biologist to analyze.	\\	\\
The woman put the cheese that she bought at the store on the broccoli that she was preparing for dinner.	\\	\\
Marcos put some decals that he bought at the store on the bus over the weekend.	\\	\\
The woman wrapped the tape that she grabbed from the storage room around the cable in the electric panel.	\\	\\
The technician put the film that he purchased yesterday in the camera before the photo shoot.	\\	\\
My son put some pins that he bought yesterday on the capitals shown on the map hanging in the classroom.	\\	\\
The woman added the sweet cream that she bought from the farmer to the caramels she was making.	\\	\\
Susana showed the house that she wants to renovate to the carpenter who lives next door.	\\	\\
The young girls returned the sculptures that they found last week to the cathedral during the festivities.	\\	\\
The girl added the sugar that she found in the cupboard to her cereal after she drank her orange juice.	\\	\\
The artist applied some strange color that he had mixed to the circle on the canvas.	\\	\\
The gentleman gave the ticket that he bought for the football game to a client from his company.	\\	\\
The Indian gave the food that he had found on the floor to the cobra during the show.	\\	\\
The guest added filtered water that he got from the refrigerator to the coffee that was sitting on the table.	\\	\\
The architect sold the book that he wrote last year to some colleagues who work at another university.	\\	\\
The soldier described the beating that he saw to the colonel of his division.	\\	\\
My sister threw a tomato that she got from the refrigerator at the composer at the end of the concert.	\\	\\
My mother installed some software that she considers very useful on the computer in the office.	\\	\\
The teacher drew the lines that he considered most important on the cube during the math lesson.	\\	\\
Simon added a change that he thought of a few minutes ago to the decision made by the committee.	\\	\\
The doctor gave the pictures that he found on the floor to the detective at the police station.	\\	\\
The nuns took the quilts that they no longer used to the director of the orphanage.	\\	\\
The thief added the gun powder that he bought the other day to the dynamite to create a huge explosion.	\\	\\
The worker sent the advertisements that he selected from the catalog to the engineer in Madrid.	\\	\\
The kitten gave the mouse that he caught last night to his family during dinner.	\\	\\
The company put up a door that it manufactured last week on the garage of my neighbor's home	\\	\\
The scientist gave the plan that he had drafted to the general during the meeting.	\\	\\
The boy gave an apple that he had in his bag to the giraffe during his visit to the zoo.	\\	\\
The mother put the lettuce that she bought at the store on the hamburger before serving it.	\\	\\
The geologist sent the stones that he analyzed last month to the institute of natural history.	\\	\\
The farmer put the ointment that she brought with her on the kangaroo to make it feel better.	\\	\\
The magician told the secrets that he had gathered over the years to the members of the audience after his show.	\\	\\
The spy turned over the sketches that he found in the drawer to the official in Iran.	\\	\\
The artist sent the pictures that he finished in his studio to the organizer of the fundraiser.	\\	\\
The therapist recommended the diet that he had followed himself to the patient during the appointment.	\\	\\
The agent applied the solution that he had in his case to the pipe to find the fingerprints.	\\	\\
The prisoner gave the treasure that he had in the cupboard to the pirate during the fight on the boat.	\\	\\
The man deactivated the safety mechanism that he had installed on the pistol the day before the accident.	\\	\\
The farmer added the fertilizer that he bought at the store to the plant that was dying.	\\	\\
The chef put the piece of meat that he had stuffed the day before on the plate during the TV program.	\\	\\
The teacher gave the discussion topic that he had chosen for the conference to the presenter a month ago.	\\	\\
The diplomat described the country that she visited to the president who hadn't been there yet.	\\	\\
The business man added several issues that he was worried about to the list of problems that will be discussed tomorrow.	\\	\\
The students recounted the story that they heard the other day to the professor of English literature.	\\	\\
The linguist added the comments that she had thought of the night before to the project written by the students.	\\	\\
The girl gave a piece of bread that she had in the pantry to the rat in the apartment.	\\	\\
The woman explained the illness that she had come down with to the receptionist who scheduled her an appointment.	\\	\\
The defendant revealed the information that he found in his cell phone to the reporter during the press conference.	\\	\\
The wife put the cover that she had sewn yesterday on the sofa to hide all the stains.	\\	\\
The girl added the cheese that she liked the most to the soup that her mother prepared.	\\	\\
The man bolted the plaque that arrived yesterday to the statue of the unknown soldiers.	\\	\\
Sandra added some new ideas that she read about in a book to her strategy for increasing her success.	\\	\\
The mathematician recommended the program that she had created to the best student in her class.	\\	\\
My mother sewed some silk flowers that she had designed to the sweater that she was knitting for my sister.	\\	\\
The specialist welded the replacement part that he bought at the store to the tank at the campsite.	\\	\\
The inspector connected the microphone that he bought at the store to the telephone in his house.	\\	\\
The guide showed the trail that he knew about to the tourists who were waiting to go hiking.	\\	\\
The farmers gave the apples that they harvested last week to the vendor at the market.	\\	\\
The photographer applied the varnish that he had mixed in his study to the violin in the auditorium.	\\	\\

\section{English --- Syntax Specific --- Noncognates}

The Romans used to dedicate a tribute that they organized every year to the beauty of women.	\\	\\
The priest gave a slice of bread that he had in his backpack to the beggar in the street.	\\	\\
The millionaire donated the money that he earned over his lifetime to the blind to help them.	\\	\\
The father gave the snacks that he warmed up in the microwave to the boys before the game.	\\	\\
The architect added the final details that he had dreamed of to the building a month before it opened.	\\	\\
The electrician installed the lights that he bought yesterday on the cave behind the Roman ruins.	\\	\\
The governor dedicated a tribute that he had organized with his friends to the childhood of the founder of the university.	\\	\\
The notary public put the accent marks that he saw were missing on the deed of the house.	\\	\\
The pastry cook put some flowers that she had seen in a magazine on the dessert that she prepared for the event.	\\	\\
The actress brought a diamond collar that she bought at the jewelers to her dog after her trip.	\\	\\
The man put the flyer that he had printed on the elevator before leaving the building.	\\	\\
The ogre showed the ring that he was protecting to the elves who were hiding him.	\\	\\
The man gave a lobster that he grabbed from a case to the fisherwoman so that she could weigh it.	\\	\\
The cook added the amount of water that he read off the recipe to the flour to prepare the dough.	\\	\\
The waiter ran a spoon that he had behind the bar across the foam of the beer.	\\	\\
The car rental office gave the GPS they owned to the foreigners so they wouldn't get lost.	\\	\\
The girl attached the drawing that she had done at school to the freezer in the basement.	\\	\\
The girl told the secret that she had learned last night to her friend from high school.	\\	\\
During the meeting the doctors put limits that they stipulated as necessary on the growth of the hospital.	\\	\\
The hotel concierge offered the soccer tickets that he had just received to the guests after he spilled a vase of flowers all over them.	\\	\\
The woman added the color that she had mixed to her hair while she was standing in the shower.	\\	\\
The ballerina put the lid that she had lost the day before on the hairspray after putting her make-up on.	\\	\\
The pilot contributed the experience that he acquired over many years to the handling of the supersonic jet.	\\	\\
The plumber hammered a plank that he had in his van onto the hole under the sink.	\\	\\
The man put the saddle that he designed on the horse to make sure it fit.	\\	\\
The geologist added some tin that he had melted down in his shop to the iron in order to create steel.	\\	\\
The technician put the screw that he found yesterday on the ladder of the airplane.	\\	\\
The father put the rub that he had previously mixed together on the lamb before cooking it.	\\	\\
My children asked the questions that they had been thinking of to the librarian after she was done talking.	\\	\\
The swimmer removed the fat that he didn't want to cook from the loin that he bought at the butcher's shop.	\\	\\
The painter gave a picture that he had painted during his free time to the manager who assists him.	\\	\\
The police officer restored the peace that he had sought for so long to the neighborhood with the capture of the thieves.	\\	\\
The mother glued the label that she had filled out on to the notebook that was on the table.	\\	\\
The farmer took the egg that he saw in the nest from the ostrich before leaving the farm.	\\	\\
The host of the festivities donated some antique chariots that he owned to the parade in the small town.	\\	\\
The school board added some restrictions that they felt were necessary to the party for the end of the year.	\\	\\
My mother added the custard that she had prepared to the peaches before putting them in the oven.	\\	\\
The judges awarded a special prize that they give every year to the best performance in the street.	\\	\\
The photographer added some wax that he had bought to the plums to make them look more appetizing.	\\	\\
The cook put a spicy sauce that he cooked with his friend on the potato that he was going to bake.	\\	\\
The secretary replaced the ink cartridge that she bought at the store on the printer before going home.	\\	\\
The nuns included the work that they hadn't completed to the punishment they gave to the girls.	\\	\\
The girl gave the toy that she had bought to her puppy after he sat and patiently waited for it.	\\	\\
The children showed the firewood that they had cut to the rangers who were in the cabin.	\\	\\
The soccer organization gave the award that it had created to the referee for his hard work.	\\	\\
The accountant made some corrections that he considered necessary to the report that his associate gave him.	\\	\\
The seamstress put some decorations that she had designed on the ribbon of the wedding dress.	\\	\\
The architect added the new base that he designed to the scale before showing it to his boss.	\\	\\
The technician installed the new rods that he received yesterday on the screen that separates his dining room from his living room.	\\	\\
The soldier sewed the patches that he got at the mall to the socks in his drawer.	\\	\\
According to the legend the devil snatched the happiness that he had longed for from the soul that was rising to heaven.	\\	\\
The instructor poured a few drops of water that she had in a bottle on the spark of the cigarette.	\\	\\
The woman attributed the cause of the allergies that she had on spring and its flowers.	\\	\\
The man added some spices that he bought in the market to the steak before cooking it.	\\	\\
The man threw a chair that he grabbed from the dining room at his stepbrother and then ran away.	\\	\\
The teacher erased the questions that she hadn't written from the survey for her students.	\\	\\
The commentator removed the wine stain that he noticed from his tie before the show.	\\	\\
The architect glued some marble pieces that he had sanded down on to the tower in order to restore it.	\\	\\
The gypsy sold a sprig of rosemary that he cut from the tree to the traveler that was walking around the square.	\\	\\
The man added some bricks that he collected yesterday to the wall of the house he was building.	\\	\\
The woman added the shoe laces that she had on her shoes to the wash that she was doing for her daughter.	\\	\\
The man put the furniture that he bought at the store yesterday in his workroom on the first floor.	\\	\\
The saleswoman applied a disinfectant that she had in her bag to the wound that she got while moving some boxes.	\\	\\
The seamstress wanted to sew some small flowers that she had bought on to the zipper of her daughter's jacket.	\\	\\

\section{Spanish --- Syntax Nonspecific --- Cognates}
El le\'{o}n que estaba sentado en el campo fij\'{o} la mirada en el ant\'{i}lope mientras com\'{i}a hierba junto la laguna.	\\	\\
El granjero que trabajaba intensamente en el campo odiaba al artista por la vida tan tranquila que llevaba.	\\	\\
Los profesores que fueron al evento animaron al atleta durante su carrera final.	\\	\\
La mujer que hab\'{i}a hecho la llamada a la polic\'{i}a habl\'{o} con las autoridades acerca del robo cuando llegaron.	\\	\\
La panadera que trabaja cerca de la iglesia tiene un bebé de cuatro a\~{n}os que ya puede tocar el piano.	\\	\\
La muchucha que conoc\'{i} en la tienda es bi\'{o}loga y fabrica productos farmacéuticos.	\\	\\
La chica que siempre olvida d\'{o}nde deja las llaves fue a la tienda a comprar br\'{o}coli para el platillo que estaba preparando.	\\	\\
El hombre que estaba regresando de la cena condujo el bus con los maestros de la escuela.	\\	\\
El cartero que trabaja para la ciudad cort\'{o} el cable que colgaba de la ventada porque representaba un peligro.	\\	\\
El ladr\'{o}n que rob\'{o} el banco bebi\'{o} un café antes del atraco y luego le dijo a sus amigos lo que hizo.	\\	\\
El abogado que estudi\'{o} en Harvard pidi\'{o} la c\'{a}mara por internet porque era m\'{a}s barata.	\\	\\
Mi pap\'{a} quien ha viajado por el mundo no hab\'{i}a visto un canguro hasta su viaje a Australia .	\\	\\
El historiador que estaba sentado en el parque sab\'{i}a todas las capitales de los pa\'{i}ses africanos.	\\	\\
La psic\'{o}loga que estudia la memoria adora cuando sus participantes traen caramelos como obsequio.	\\	\\
El mayordomo que trabaja en la mansi\'{o}n llam\'{o} al carpintero para reparar la puerta principal de la casa de visitas.	\\	\\
El cantero que era primo del rey construy\'{o} una catedral con material reciclado.	\\	\\
El d\'{a}lmata que ten\'{i}a las orejas ca\'{i}das se comi\'{o} el cereal que el ni\~{n}o derram\'{o} al piso.	\\	\\
El ni\~{n}o que est\'{a} en tercer grado dibuj\'{o} un c\'{i}rculo con su l\'{a}piz azul y luego lo cort\'{o} y lo peg\'{o} en su libro	\\	\\
La mujer que hab\'{i}a trabajado en una empresa que hace papel vio a su cliente en el centro comercial.	\\	\\
La actriz que estaba en el balc\'{o}n mat\'{o} a una cobra con una pala que sac\'{o} del dep\'{o}sito	\\	\\
El librero que abri\'{o} la nueva tienda en la esquina invit\'{o} a sus colegas a la inauguraci\'{o}n.	\\	\\
La actriz que estudi\'{o} en Madrid cen\'{o} con el compositor de San Francisco.	\\	\\
La cajera que trabaja en el banco quem\'{o} la computadora de su escritorio con su cigarrillo.	\\	\\
El telegrafista que trabaja en la base aérea avis\'{o} al coronel del ataque que estaban planeando los enemigos.	\\	\\
El asistente que ense\~{n}a en la escuela dise\~{n}\'{o} el cubo para la clase de f\'{i}sica.	\\	\\
El congresista que vive en Minnesota argument\'{o} que su decisi\'{o}n fue la correcta dadas las circunstancias.	\\	\\
El escritor que estaba tratando de buscar ideas para su novela escuch\'{o} al detective durante el interrogatorio.	\\	\\
El soldado que estaba ansioso por demostrar sus capacidades encendi\'{o} la dinamita demasiado pronto y casi pone la misi\'{o}n en peligro.	\\	\\
El  empresario que hab\'{i}a invertido mucho dinero despidi\'{o} al director de la compa\~{n}\'{i}a el mes pasado.	\\	\\
La bailarina que estaba en la exposici\'{o}n comprar\'{a} una estatua en Berl\'{i}n durante su pr\'{o}xima visita.	\\	\\
El entrenador que trabaja en la escuela cree que su estrategia dar\'{a} buenos resultados.	\\	\\
La mujer que asisti\'{o} a un campamento de verano de ni\~{n}a se convirti\'{o} en la mejor estudiante de su clase.	\\	\\
El hombre que ama a los animales llev\'{o} a su casa un gatito para su familia sin antes decirle a su esposa.	\\	\\
La peluquera que compr\'{o} un coche abri\'{o} el garaje de su casa con su mando.	\\	\\
El rey que gobernaba el pa\'{i}s invit\'{o} al general a comer gambas y tomar una cerveza.	\\	\\
El jardinero que fue contratado por la escuela prepar\'{o} una hamburguesa para comer con sus amigos.	\\	\\
El contable que tiene tres hijos llam\'{o} al ingeniero el d\'{i}a de su boda para felicitarlo.	\\	\\
El arque\'{o}logo que ha dedicado su vida a las ciencias inaugur\'{o} el instituto de Investigaci\'{o}n Anat\'{o}mica.	\\	\\
Los bomberos que llegaron al sitio del accidente ayudaron a la jirafa en peligro en el zoo de San Diego.	\\	\\
Las personas que fundaron las Naciones Unidas eran miembros de una generaci\'{o}n esperanzada de individuos.	\\	\\
El rehén que estaba atrapado en el banco llam\'{o} al oficial miestras el ladr\'{o}n estaba distra\'{i}do.	\\	\\
La enfermera que trabajar\'{a} con el cardi\'{o}logo empuj\'{o} al organizador por la puerta.	\\	\\
El especialista que es muy famoso habl\'{o} con el paciente sobre varios tratamientos para curar su enfermedad.	\\	\\
El dentista que muri\'{o} la semana pasada hab\'{i}a comprado una pipa en Turqu\'{i}a para su colecci\'{o}n.	\\	\\
El hombre que navegaba por los mares en su barco mat\'{o} al pirata con un cuchillo.	\\	\\
La camarera que fuma dos cajetillas de cigarrillos al d\'{i}a carga una pistola en su bolso.	\\	\\
El pol\'{i}tico que vive en Washington compr\'{o} una planta para su esposa porque le gustan mucho.	\\	\\
La mujer que abri\'{o} la vitrina tir\'{o} un plato a su marido porque estaba molesta con él.	\\	\\
El narrador de noticias que hab\'{i}a sido ascendido recientemente anim\'{o} al presentador a actuar con naturalidad en el aire.	\\	\\
La diplom\'{a}tica que viaja a muchos pa\'{i}ses pregunt\'{o} al presidente si pod\'{i}a convertirse en una esp\'{i}a.	\\	\\
El decorador que escribe para la revista predijo que los problemas con la pintura ser\'{i}an dif\'{i}cil de resolver.	\\	\\
El  taxista que estaba estacionado en la esquina de la panader\'{i}a  llev\'{o} a la profesora a su casa.	\\	\\
El explorador que fue entrevistado ayer dijo que el proyecto requer\'{i}a m\'{a}s dinero para que se puediera completar.	\\	\\
El se\~{n}or que organiz\'{o} el evento encontr\'{o} una rata en la despensa de la cocina la semana pasada.	\\	\\
La hermana del gobernador que lo ayud\'{o} durante su campa\~{n}a recogi\'{o} las cartas de la recepci\'{o}nista antes de irse.	\\	\\
El esquiador que gan\'{o} la medalla en las olimpiadas vio al reportero durante el salto.	\\	\\
La pelota de béisbol que fue autografada por el jugador famoso aterriz\'{o} en el sof\'{a} después de haberse caido de la vitrina.	\\	\\
La se\~{n}ora que vive en nuestro edificio dej\'{o} la sopa en la ventana para que se enfriara.	\\	\\
El soldado que hab\'{i}a luchado en la batalla cosi\'{o} el suéter en su tienda de campa\~{n}a.	\\	\\
El pacifista que hab\'{i}a organizado la manifestaci\'{o}n pint\'{o} el tanque con s\'{i}mbolos de paz.	\\	\\
La cuidadora que empez\'{o} a trabajar ayer limpi\'{o} el teléfono con agua y jab\'{o}n para desinfectarlo.	\\	\\
El esp\'{i}a que facilit\'{o} los datos interrog\'{o} a los turistas en Canc\'{u}n para obtener m\'{a}s informaci\'{o}n sobre el caso.	\\	\\
Los chicos que estaban en el supermercado llamaron al vendedor por la ventana.	\\	\\
El actor que vivi\'{o} en Chile compr\'{o} un viol\'{i}n para el cumplea\~{n}os de su hija en una tienda famosa.	\\	\\

\section{Spanish --- Syntax Nonspecific --- Noncognates}
La ni\~{n}a que quiere ser artista recibi\'{o} un premio por su actuaci\'{o}n en el acto escolar.	\\	\\
El astr\'{o}nomo que trabajaba en el observatorio en Houston ten\'{i}a un agujero en la chaqueta.	\\	\\
El cura que da la misa los domingos salv\'{o} el alma de todos los pecadores la semana pasada.	\\	\\
La mujer que fue contratada por la compa\~{n}ia era amiga del ejecutivo que vivi\'{o} en Panam\'{a}.	\\	\\
El hombre que disfruta cuando hace ejercicio decidi\'{o} convertirse en arbitro después de haber visto un partido de f\'{u}tbol.	\\	\\
El agente que odiaba la actividad f\'{i}sica llam\'{o} el ascensor para ir al primer piso.	\\	\\
La gu\'{i}a que visit\'{o} varias ciudades en Australia prepar\'{o} un avestruz en la cocina.	\\	\\
El vecino que viv\'{i}a en el sexto piso del edificio adorn\'{o} el barrio para los festejos.	\\	\\
La enfermera que ten\'{i}a una hora libre después de su almuerzo limpi\'{o} la b\'{a}scula en el laboratorio.	\\	\\
El p\'{u}blico que asisti\'{o} al desfile de moda reconoci\'{o} la belleza de la modelo de Ruman\'{i}a.	\\	\\
La mujer que necesitaba un libro detuvo al bibliotecario para hacerle una pregunta.	\\	\\
La maestra que quer\'{i}a decorar la habitaci\'{o}n de su nueva casa compr\'{o} un biombo el viernes.	\\	\\
El fontanero que ten\'{i}a tres hijos en la universidad prepar\'{o} un bistec con verdura.	\\	\\
El portero que jugaba al f\'{u}tbol en el colegio cosi\'{o} la bragueta de los pantalones antes del partido.	\\	\\
El escritor que lleg\'{o} ayer a Madrid vio la cabalgata desde el balc\'{o}n con sus amigos.	\\	\\
El chico que quiere ser veterinario quer\'{i}a comprar un caballo con el dinero de su cuenta de ahorros.	\\	\\
La ni\~{n}a que es gimnasta ol\'{i}mpica no quiere cortarse la cabellera a pesar de que est\'{a} muy larga.	\\	\\
Mi hijo que come carne dijo que su amiga compr\'{o} cabrito para el picnic.	\\	\\
La zanahoria que el granjero recogi\'{o} ese d\'{i}a golpe\'{o} al cachorro haciéndolo ge\~{n}ir.	\\	\\
La limpiadora que necesitaba ropa para su trabajo pidi\'{o} unos calcetines por internet.	\\	\\
La nadadora que comi\'{o} dulces y galletas recibi\'{o} un castigo de la entrenadora antes de la competici\'{o}n.	\\	\\
El hombre que hab\'{i}a tomado un curso sobre supervivencia cre\'{o} la chispa para empezar el fuego.	\\	\\
El perro que era muy dulce cuidaba a los ciegos para mantenerlos fuera de peligro.	\\	\\
La hermana que es muy meticulosa compr\'{o} una cinta para el velo de la novia.	\\	\\
El hombre que prepara mermeladas hizo un puré de ciruelas y lo puso a cocinar.	\\	\\
La abuela que hab\'{i}a ganado la loter\'{i}a compr\'{o} un congelador para su casa nueva.	\\	\\
El hombre de negocios que vest\'{i}a muy bien compr\'{o} una corbata en la tienda al lado de su casa.	\\	\\
El zo\'{o}logo que hac\'{i}a experimentos con monos fren\'{o} su crecimiento con el compuesto que hab\'{i}a preparado.	\\	\\
El chico que vive cerca de mi casa compr\'{o} un cuaderno para tomar apuntes en un paseo.	\\	\\
El historiador que estaba grabando un programa de TV investig\'{o} la cueva antes de empezar la grabaci\'{o}n.	\\	\\
El hombre que es muy inteligente pero desordenado limpi\'{o} el despacho antes de que llegara su amigo.	\\	\\
La historia que el ni\~{n}o estaba leyendo involucraba un grupo de duendes que estaban tratando de salvar el mundo de los gnomos malos.	\\	\\
La peque\~{n}a ni\~{n}a que llevaba su vestido favorito comi\'{o} duraznos en el auto de camino a casa.	\\	\\
El hombre de negocios que trabajaba arduamente vi\'{o} el edificio mientras lo estaban demoliendo.	\\	\\
El ministro que volvi\'{o} de sus vacaciones en Italia despidi\'{o} al encargado por haber sido incompetente.	\\	\\
El estudiante que particip\'{o} en un experimento devolvi\'{o} la encuesta para el estudio en un sobre.	\\	\\
El pintor que trabajaba para mi padre por las ma\~{n}anas sac\'{o} la escalerilla de la camioneta.	\\	\\
El abogado que tramitaba la compra del edificio mand\'{o} la escritura al ayuntamiento desde la oficina de correos.	\\	\\
La se\~{n}ora que estaba lavando los platos se quit\'{o} la espuma de las manos antes de contestar el puerta	\\	\\
El pa\'{i}s que fue formado hace diez a\~{n}os prohibi\'{o} que los extranjeros se convirtieran en residentes.	\\	\\
El escritor que estaba concentrado en su nuevo libro no oy\'{o} la fiesta en el parque.	\\	\\
El entrenado que gritaba con fuerza durante el partido rompi\'{o} el folio con sus anotaciones.	\\	\\
El oficial de policia que visit\'{o} el parque arrest\'{o} al guardabosques que era sospechoso de robar un banco.	\\	\\
El tartero que ten\'{i}a una pasteler\'{i}a en la ciudad tir\'{o} la harina por la calle principal para protestar.	\\	\\
El luchador que derrot\'{o} al campe\'{o}n del mundo cubri\'{o} su herida con una venda después de la pelea.	\\	\\
El médico que bebi\'{o} demasiadas cervezas en el bar hiri\'{o} a su hermanastro en una pelea.	\\	\\
El jinete que gan\'{o} la ultima carrera del Kentucky Derby compr\'{o} hierro para su establo.	\\	\\
La mujer que mantiene su casa muy limpia persuadi\'{o} los huéspedes a que se quitaran los zapatos antes de entrar.	\\	\\
El representante que alquil\'{o} un coche en el aeropuerto arregl\'{o} la impresora para la empresa.	\\	\\
La estilista que tiene una tienda en el centro vende laca y otros productos para el cabello.	\\	\\
El estudiante que ten\'{i}a muchos quehaceres complet\'{o} el lavado con suavizante antes de volver a casa.	\\	\\
El piloto que dio una celebraci\'{o}n en su casa compr\'{o} el mejor lomo y jam\'{o}n que vend\'{i}an en la tienda.	\\	\\
La ayudante que est\'{a} encargada de preparar las pizzas aprendi\'{o} el correcto manejo de la m\'{a}quina para hacer masas.	\\	\\
La monja que hab\'{i}a perdido la cabeza atac\'{o} al mendigo en la cocina durante Acci\'{o}n de Gracias.	\\	\\
El  jurado que hab\'{i}a escuchado el caso por tres d\'{i}as consecutivos miraba hacia el muro cuando declar\'{o} que necesitaba un descanso.	\\	\\
El bombero que estaba en peligro dentro del edificio record\'{o} su ni\~{n}ez al salvar al recién nacido.	\\	\\
El hombre que vive por aqu\'{i} cerca tiene tres hijos que lo siguen por d\'{o}nde quiera que va.	\\	\\
La cocinera que prerara guisos compr\'{o} una papa en el mercado para hacer la cena.	\\	\\
El inquilino que ama los animales siempre peina a su perro con un cepillo de cerda natural.	\\	\\
El conductor que ten\'{i}a el d\'{i}a libre por las vacaciones llam\'{o} a la pescadora para salir.	\\	\\
El dietista que celebr\'{o} su cumplea\~{n}os con sus amigos prepar\'{o} un postre para la cena.	\\	\\
El adolescente que estaba enamorado de una amiga esper\'{o} la llegada de la primavera para declararle su amor.	\\	\\
El camarero que corr\'{i}a todas las ma\~{n}anas lleg\'{o} a la torre en menos de una hora.	\\	\\
La actriz que amaba su trabajo m\'{a}s que nada invit\'{o} al viajero a una copa de vino.	\\	\\

\section{Spanish Sentences --- Syntax Specific --- Cognates}
La veterinaria le inyect\'{o} el antibi\'{o}tico que ten\'{i}a en el gabinete al ant\'{i}lope cuando empez\'{o} a temblar de forma incontrolable.	\\	\\
El hombre rico le dio la fotograf\'{i}a que hab\'{i}a tomado de su esposa al artista para que pintara su retrato.	\\	\\
La enfermera le dio la medicina que disolvi\'{o} en agua a la atleta con dolor de cabeza.	\\	\\
El sospechoso le dio el cuchillo que hab\'{i}a estado escondiendo a las autoridades cuando decidi\'{o} entregrarse.	\\	\\
El t\'{i}o le regal\'{o} la cuna que hab\'{i}a restaurado hac\'{i}a unos meses al bebé para su cumplea\~{n}os.	\\	\\
La polic\'{i}a le dio la muestra de sangre que encontr\'{o} en la escena del cr\'{i}men a la bi\'{o}loga para que la analizara.	\\	\\
La mujer le puso el queso que compr\'{o} en la tienda al br\'{o}coli que estaba preparando para la cena.	\\	\\
Marcos le coloc\'{o} unas calcoman\'{i}as que compr\'{o} en la tienda al bus durante el fin  de semana.	\\	\\
La mujer le peg\'{o} la cinta aislante que us\'{o} el a\~{n}o pasado al cable en el cuadro eléctrico.	\\	\\
El invitado le agreg\'{o} el agua filtrada que sac\'{o} de la nevera al café que estaba en la mesa.	\\	\\
El inform\'{a}tico le puso el carrete que compr\'{o} ayer a la c\'{a}mara antes de la sesi\'{o}n fotogr\'{a}fica.	\\	\\
La granjera le puso la pomada que trajo con ella al canguro para hacerlo sentir mejor.	\\	\\
Mi hijo le puso unas chinchetas que compr\'{o} ayer a las capitales en el mapa colgado en la clase.	\\	\\
La mujer le a\~{n}adi\'{o} la crema dulce que compr\'{o} del granjero a los caramelos que estaba preparando.	\\	\\
Susana le mostr\'{o} la casa que quiere renovar al carpintero que vive al lado.	\\	\\
Las chicas le entregaron las esculturas que encontraron la semana pasada a la catedral durante las fiestas.	\\	\\
La ni\~{n}a le a\~{n}adi\'{o} el az\'{u}car que hab\'{i}a encontrado en el aparador al cereal después de haber tomado el zumo de naranja.	\\	\\
El artista le aplic\'{o} un color extra\~{n}o que hab\'{i}a mezclado al c\'{i}rculo en el lienzo.	\\	\\
El se\~{n}or le regal\'{o} la entrada que compr\'{o} para el juego de futbol a un cliente de su oficina.	\\	\\
El indio le dio la comida que hab\'{i}a encontrado en el suelo a la cobra durante el espect\'{a}culo.	\\	\\
El arquitecto les vendi\'{o} el libro que escribi\'{o} el a\~{n}o pasado a unos colegas que trabajan en otra universidad.	\\	\\
Mi hermana le lanz\'{o} un tomate que cogi\'{o} del frigor\'{i}fico al compositor al final del concierto.	\\	\\
Mi madre le instal\'{o} un programa que considera muy \'{u}til a la computadora en la oficina.	\\	\\
El soldado le describi\'{o} la paliza que vi\'{o} al coronel de su escuadr\'{o}n.	\\	\\
El maestro le dibuj\'{o} las l\'{i}neas que consideraba m\'{a}s importantes al cubo durante la lecci\'{o}n de matem\'{a}ticas.	\\	\\
Sim\'{o}n le a\~{n}adi\'{o} un cambio que se le occuri\'{o} hace pocos minutos a la decisi\'{o}n de la junta.	\\	\\
El doctor le dio las fotograf\'{i}as que encontr\'{o} en el suelo al detective en la comisar\'{i}a.	\\	\\
El ladr\'{o}n le a\~{n}adi\'{o} la p\'{o}lvora que hab\'{i}a comprado el otro d\'{i}a a la dinamita para crear una explosi\'{o}n enorme.	\\	\\
Las monjas le llevaron las mantas que no usaban m\'{a}s al director del orfanato.	\\	\\
El hombre le atornill\'{o} la placa que lleg\'{o} ayer a la estatua de los soldado caidos.	\\	\\
Sandra le a\~{n}adi\'{o} algunas ideas que ley\'{o} en un libro a su estrategia para incrementar su éxito.	\\	\\
La matem\'{a}tico le recomend\'{o} el programa que hab\'{i}a escrito al mejor estudiante de su clase.	\\	\\
El gatito le di\'{o} el rat\'{o}n que atrap\'{o} anoche a su familia durante la cena.	\\	\\
La compa\~{n}\'{i}a le coloc\'{o} una puerta que manufacturaron la semana pasada al garaje de la casa de mis vecinos.	\\	\\
El cient\'{i}fico le dio el plan que hab\'{i}a ideado al general durante la reuni\'{o}n.	\\	\\
La madre le puso la lechuga que compr\'{o} en la tienda a la hamburguesa antes de servirla.	\\	\\
El obrero le envi\'{o} los anuncios que seleccion\'{o} del cat\'{a}logo al ingeniero en Madrid.	\\	\\
El ge\'{o}logo le mand\'{o} las piedras que analiz\'{o} el mes pasado al instituto de historia natural.	\\	\\
El ni\~{n}o le dio una manzana que ten\'{i}a en su bolso a la jirafa durante la visita al zoo.	\\	\\
El mago le dijo algunos secretos que hab\'{i}a compilado a lo largo de los a\~{n}os a los miembros de la audiencia después de su demonstraci\'{o}n.	\\	\\
El esp\'{i}a le facilit\'{o} los planos que encontr\'{o} en el caj\'{o}n al oficial en Ir\'{a}n.	\\	\\
El artista le envi\'{o} los cuadros que termin\'{o} en su estudio al organizador del evento benéfico.	\\	\\
El terapeuta le recomend\'{o} la dieta que hab\'{i}a seguido a la paciente durante la consulta.	\\	\\
El agente le aplic\'{o} la soluci\'{o}n que ten\'{i}a en su malet\'{i}n a la pipa para encontrar las huellas..	\\	\\
El prisionero le dio el tesoro que ten\'{i}a en el armario al pirata durante la pelea en el barco.	\\	\\
El hombre le quit\'{o} el seguro que hab\'{i}a instalado a la pistola el d\'{i}a antes del accidente.	\\	\\
El agricultor le a\~{n}adi\'{o} el fertilizante que compr\'{o} en la tienda a la planta que se estaba muriendo.	\\	\\
El cocinero le puso el trozo de carne que hab\'{i}a rellenado el d\'{i}a anterior al plato durante el programa de televisi\'{o}n.	\\	\\
La maestra le dio el tema de la discusi\'{o}n que hab\'{i}a elegido para la conferencia al presentador hace un mes.	\\	\\
La diplom\'{a}tica le describi\'{o} el pa\'{i}s que visit\'{o} al presidente que todav\'{i}a no hab\'{i}a estado ah\'{i}.	\\	\\
El empresario le agreg\'{o} varios temas que le preocupan a la lista de problemas que se discutir\'{a} ma\~{n}ana.	\\	\\
Los estudiantes le contaron el cuento que oyeron el otro d\'{i}a a la profesora de literatura inglesa.	\\	\\
El ling\:{u}ista le a\~{n}adi\'{o} los comentarios que hab\'{i}a pensado la noche anterior al proyecto escrito por los estudiantes.	\\	\\
La ni\~{n}a le dio un trozo de pan que ten\'{i}a en la despensa a la rata en el apartamento.	\\	\\
La se\~{n}ora le explic\'{o} la enfermedad que ten\'{i}a a la recepcionista que le dio una cita.	\\	\\
El defensor le revel\'{o} la informaci\'{o}n que encontr\'{o} en el m\'{o}vil al reportero durante la rueda de prensa.	\\	\\
La esposa le puso la funda que hab\'{i}a cosido ayer  al sof\'{a} para esconder las manchas.	\\	\\
La ni\~{n}a le a\~{n}adi\'{o} el queso que m\'{a}s le gustaba a la sopa que prepar\'{o} su mam\'{a}.	\\	\\
Mi madre le cosi\'{o} unas flores de seda que hab\'{i}a dise\~{n}ado al suéter que estaba tejiendo para mi hermana.	\\	\\
El especialista le sold\'{o} el repuesto que compr\'{o} en la tienda al tanque en el campamento.	\\	\\
El inspector le conect\'{o} el micr\'{o}fono que compr\'{o} en la tienda al teléfono de su casa.	\\	\\
El gu\'{i}a le mostr\'{o} el camino que conoc\'{i}a a los turistas que estaban esperando para hacer senderismo.	\\	\\
Los agricultores le dieron las manzanas que cosecharon la semana pasada al vendedor en el mercado.	\\	\\
El fot\'{o}grafo le aplic\'{o} el barniz que hab\'{i}a mezclado en su estudio al viol\'{i}n en el auditorio.	\\	\\

\section{Spanish --- Syntax Specific --- Noncognates}

Los jueces le otorgaron un premio especial que dan cada a\~{n}o a la mejor actuaci\'{o}n en la calle.	\\	\\
El fontanero le clav\'{o} una tabla que consigui\'{o} en su camioneta al agujero debajo del lavamanos.	\\	\\
Seg\'{u}n la leyenda el demonio le arrebat\'{o} la felicidad que tanto deseaba al alma que sub\'{i}a al cielo.	\\	\\
La muchacha le cont\'{o} el secreto que hab\'{i}a guardado a su amiga de la escuela.	\\	\\
La organizaci\'{o}n de f\'{u}tbol le dio el premio que hab\'{i}a creado al \'{a}rbitro por su arduo trabajo.	\\	\\
El hombre le coloc\'{o} el anuncio que imprimi\'{o} al ascensor antes de irse del edificio.	\\	\\
El granjero le quit\'{o} el huevo que vio en el nido a la avestruz antes de irse de la granja.	\\	\\
El polic\'{i}a le devolvi\'{o} la paz que hab\'{i}a buscado por tanto tiempo al barrio con la captura de los ladrones.	\\	\\
El arquitecto le puso la base que dise\~{n}\'{o} a la b\'{a}scula antes de mostrasela a su jefe.	\\	\\
Los romanos le dedicaban un tributo que organizaban todos los a\~{n}os a la belleza de la mujer.	\\	\\
Mis hijos le hicieron las preguntas que hab\'{i}an estado pensando al bibliotecario después de que él termin\'{o} de hablar.	\\	\\
El técnico le clav\'{o} las varas nuevas que recibi\'{o} ayer al biombo que separa el comedor de la sala.	\\	\\
El hombre le a\~{n}adi\'{o} unas especias que compr\'{o} en el mercado al bistec antes de cocinarlo.	\\	\\
La costurera le quer\'{i}a coser unas florecillas que hab\'{i}a comprado a la bragueta de la chaqueta de su hija.	\\	\\
El anfitri\'{o}n del evento festivo le don\'{o} algunas carrozas antiguas que ten\'{i}a a la cabalgata del pueblito.	\\	\\
El hombre le puso la silla de montar que dise\~{n}\'{o} al caballo para ajustarla.	\\	\\
La mujer le agreg\'{o} el color que hab\'{i}a preparado a su cabellera mientras estaba en la ducha.	\\	\\
El pap\'{a} le puso las especias que hab\'{i}a mezclado previamente al cabrito antes de cocinarlo.	\\	\\
La ni\~{n}a le dio el juguete que hab\'{i}a comprado a su cachorro despu\'{e}s de que \'{e}l se sent\'{o} y esper\'{o} pacientemente.	\\	\\
El soldado le cosi\'{o} los parches que hab\'{i}a comprado en el centro comercial a los calcetines en el caj\'{o}n.	\\	\\
Las monjas le incluyeron el trabajo que no hab\'{i}an completado al castigo de las chicas.	\\	\\
La instructora le tir\'{o} unas gotas de agua que ten\'{i}a en una botella a la chispa del cigarillo.	\\	\\
El millonario le dio el dinero que hab\'{i}a estado ahorrado a los ciegos para ayudarlos.	\\	\\
La modista le puso unos adornos que hab\'{i}a dise\~{n}ado a la cinta del vestido de novia.	\\	\\
El fot\'{o}grafo le a\~{n}adi\'{o} un poco de brillantina que hab\'{i}a comprado a las ciruela para hacerlas ver m\'{a}s apetitosas.	\\	\\
La ni\~{n}a le peg\'{o} el dibujo que hab\'{i}a hecho en el colegio al congelador en el s\'{o}tano.	\\	\\
El comentarista le quit\'{o} la mancha de vino que not\'{o} a la corbata antes del programa.	\\	\\
Durante la reuni\'{o}n los m\'{e}dicos le pusieron los l\'{i}mites que estipularon necesarios al crecimiento del hospital.	\\	\\
La mam\'{a} le peg\'{o} la etiqueta que hab\'{i}a completado al cuaderno que estaba en la mesa.	\\	\\
El eletricista le instal\'{o} las luces que compr\'{o} ayer a la cueva detr\'{a}s de las ruina romanas.	\\	\\
El hombre le puso los muebles que compr\'{o} en la tienda ayer en su despacho del primer piso.	\\	\\
El ogro mostr\'{o} el anillo que estaba protegiendo a los duendes que lo estaban escondiendo.	\\	\\
Mi madre le a\~{n}adi\'{o} la crema pastelera que hab\'{i}a preparado a los duraznos antes de ponerlos en el horno.	\\	\\
El arquitecto le a\~{n}adio los \'{u}ltimos detalles que hab\'{i}a so\~{n}ado al edificio un mes antes de que lo abrieran.	\\	\\
El pintor le regal\'{o} un cuadro que hab\'{i}a pintado en su tiempo libre al encargado que lo ayuda.	\\	\\
La maestra le borr\'{o} las preguntas que no hab\'{i}a escrito a la encuesta para los estudiantes.	\\	\\
El t\'{e}cnico le puso el tornillo que encontr\'{o} ayer a la escalerilla del avi\'{o}n.	\\	\\
El escribano le puso los acentos que hab\'{i}a visto que faltaban a la escritura de la casa.	\\	\\
El camarero le pas\'{o} una cuchara que ten\'{i}a detr\'{a}s de la barra a la espuma de la cerveza.	\\	\\
La oficina de alquiler de autos le di\'{o} un GPS que ten\'{i}a a los extranjeros para que no se fueran a perder.	\\	\\
La junta directiva le agreg\'{o} algunas restricciones que consider\'{o} importartes a la fiesta de fin de a\~{n}o.	\\	\\
El contable le hizo unas correcciones que consider\'{o} necesarias al folio que le entreg\'{o} su socio.	\\	\\
Los ni\~{n}os le ense\~{n}aron la le\~{n}a que hab\'{i}an cortado al guardabosques que estaba en la caba\~{n}a.	\\	\\
El cocinero le a\~{n}adi\'{o} la cantidad de agua que ley\'{o} en la receta a la harina para preparar la masa.	\\	\\
La vendedora le aplic\'{o} un desinfectante que ten\'{i}a en su bolso a la herida que se hizo moviendo unas cajas.	\\	\\
El hombre le tir\'{o} una silla que cogi\'{o} en el comedor a su hermanastro y despu\'{e}s se escap\'{o}.	\\	\\
El ge\'{o}logo le a\~{n}adi\'{o} el esta\~{n}o que hab\'{i}a fundido en su taller al hierro para conseguir acero.	\\	\\
El conserje del hotel le ofreci\'{o} las entradas de f\'{u}tbol que hab\'{i}a recibido recientemente a los hu\'{e}spedes despu\'{e}s de derramar el  jarr\'{o}n con agua encima de ellos.	\\	\\
La secretaria le cambi\'{o} el cartucho que compr\'{o} en la tienda a la impresora antes de irse a su casa.	\\	\\
La bailarina le puso la tapa que hab\'{i}a perdido el d\'{i}a anterior a la laca despu\'{e}s de maquillarse.	\\	\\
La se\~{n}ora le agreg\'{o} los cordones que ten\'{i}a en sus zapatillas al lavado de la ropa de su hija.	\\	\\
El nadador le quit\'{o} la grasa que no quer\'{i}a cocinar al lomo que compr\'{o} en la carnicer\'{i}a.	\\	\\
El piloto le a\~{n}adi\'{o} la experiencia que hab\'{i}a adquirido durante muchos a\~{n}os al manejo del avi\'{o}n supers\'{o}nico.	\\	\\
El cura le dio un trozo de pan que ten\'{i}a en su mochila al mendigo en la calle.	\\	\\
El hombre le a\~{n}adi\'{o} algunos ladrillos que recogi\'{o} ayer al muro de la casa que estaba construyendo.	\\	\\
El governador le dedic\'{o} un tributo que hab\'{i}a organizado con sus amigos a la ni\~{n}ez del fundador de la universidad.	\\	\\
El pap\'{a} le dio la merienda que calent\'{o} en el microondas a los ni\~{n}os antes del partido.	\\	\\
El cocinero le puso una salsa picante que cocin\'{o} con su amigo a la papa que iba a hornear.	\\	\\
La actriz le trajo un collar con diamantes que compr\'{o} en la joyer\'{i}a a su perro despu\'{e}s de su viaje.	\\	\\
El hombre le di\'{o} una langosta que sac\'{o} de una caja a la pescadora para que la pesara.	\\	\\
La repostera le coloc\'{o} unas flores que hab\'{i}a visto en una revista al postre que prepar\'{o} para el evento.	\\	\\
La mujer le attribuy\'{o} la causa de las alerg\'{i}as que tiene a la primavera y sus flores.	\\	\\
El architecto le peg\'{o} unas piezas de marmol que hab\'{i}a lijado a la torre para poder restaurarla.	\\	\\
El gitano le vendi\'{o} una ramita de romero que cort\'{o} del \'{a}rbol al viajero que paseaba por la plaza.
